---
title: "Synthesize/Impute Data"
author: "Kyle M. Lang"
date: "2021-10-22"
output: 
 prettydoc::html_pretty:
    theme: architect
    highlight: github
    toc: true
    #toc_float: true
    toc_depth: 2
    #numbered_sections: true
    df_print: paged
    self_contained: false
    css: my_style.css
editor_options: 
  chunk_output_type: console
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      fig.align = "center",  
                      dev.args = list(bg = "transparent")
                      )
```

In this script, we'll synthesize/impute the cleaned burn data. We have a whole 
bunch of missing data on top of the variables we need to synthesize. We'll kill
two birds with one stone by doing the imputation and synthesis simultaneously
using the **mice** package.

---

# Data Ingest

---

First, we'll clear the workspace, load necessary packages, and define some 
global variables.

```{r, warning = FALSE, message = FALSE}
rm(list = ls(all = TRUE))

library(ggplot2)
library(plotly)
library(viridis)
library(haven)
library(mice)
library(xtable)
library(knitr)

source("select_preds.R")

dataDir <- "../../data/"
plotDir <- "../figures/"
fn      <- "cleaned_burn_data.rds"
```

Now, we'll read in the data.

```{r, cache = TRUE}
dat0 <- readRDS(paste0(dataDir, fn))

dim(dat0)
head(dat0)
```

That looks to have been successful. Let's exclude two variables based on the 
findings during data cleaning.

1. The `partner` variable, because it's completely broken.
1. The `empty_partner_data` variable because it's perfectly redundant with 
`parnter_incl`.

```{r}
dat0 <- dat0[setdiff(colnames(dat0), c("partner", "empty_partner_data"))]
```

The `mice()` function won't work with labelled vectors from **haven**, so we 
need to covert our data into a normal data frame without fancy vector types.

```{r}
dat1        <- dat0
class(dat1) <- "data.frame"

facFlag <- sapply(dat1, is.factor)

dat1[!facFlag] <- lapply(dat1[!facFlag], as.numeric)
```

---

# Imputation/Synthesis

---

First, we'll try the imputation using all the data. If this doesn't work, we'll 
try excluding some of the problematic cases we discovered during data cleaning.

## Setup

We have way too many variables to use them all as predictors in the imputation 
models.

- Our first step will be to create a more parsimonious predictor matrix using 
`mice::quickpred()`.
   - We'll select all variables that have a minimum correlation of *r* = 0.3 as 
   predictors.
   - We'll use Spearmans $\rho$ because the items are all ordinal or discrete.
   - We'll exclude the ID variable, for obvious reasons.
   
```{r}
predMat <- 
  quickpred(dat1, mincor = 0.3, method = "spearman", exclude = "ID_random")
```

Let's see how many predictors we selected for each target

```{r}
pm     <- colMeans(is.na(dat1))
nPreds <- rowSums(predMat[pm > 0, ])

range(nPreds)
hist(nPreds)
```

Well, that appraoch didn't really work. A few variables only got a small number 
of predictors (the fewest being `r min(nPreds)`), but many (i.e., 
`r sum(nPreds > 100)`) got more than a 100 predictors.

Let's try something different. This is going to be stupidly inefficient, but 
let's go for it.

First, we'll generate a range of predictor matrices with varying levels of 
minimum correlation.

```{r, cache = TRUE}
predMats <- lapply(seq(0.1, 0.9, 0.1),
                   function(x, data)
                     quickpred(data, 
                               mincor = x, 
                               method = "spearman", 
                               exclude = "ID_random"),
                   data = dat1)
```

Then, we'll count the number of predictors selected for each target in each 
predictor matrix and find which predictor matrix gives the closest to 20 
predictors for each target.

```{r}
predCounts <- sapply(predMats, rowSums)
bestMat    <- apply(abs(predCounts - 20), 1, which.min)
```

Finally, we build a single predictor matrix by selecting the appropriate row out 
of the best matrix for each target variable.

```{r}
predMat <- matrix(0, 
                  ncol(dat1), 
                  ncol(dat1), 
                  dimnames = list(colnames(dat1), colnames(dat1))
                  )

targets <- colnames(dat1)[pm > 0]
for(v in targets) predMat[v, ] <- predMats[[bestMat[v]]][v, ]
```

---

We need to define which variables to overimpute for synthesis purposes.

```{r}
synStems <- c("age",
              "gender",
              "education",
              "marital",
              "Surgeries",
              "TBSA",
              "dep")

synTargets <- 
  grep(paste(synStems, collapse = "|"), colnames(dat1), value = TRUE)
```

Now, we construct a `where` matrix to tell **mice** which values to impute.

- We need to request imputations for all missing values.
- We also want to request imputations for the observed values on the variables 
listed in `synTargets`.
   - This overimputation will be our means of synthesizing the sensitive data.
   
```{r}
## Start with a standard where matrix flagging all missing values:
whereMat <- make.where(data = dat1, keyword = "missing")

## Add the overimputation information:
whereMat[ , synTargets] <- TRUE
```

We should now be ready to run the imputation.

## Generate Imputations

```{r, eval = FALSE}
miceOut <- mice(data = dat1,
                m = 25,
                maxit = 50,
                method = "cart", 
                predictorMatrix = predMat,
                where = whereMat,
                seed = 235711)

saveRDS(miceOut, paste0(dataDir, "burn_data_mids_m25_it50_cart.rds"))
```

```{r echo = FALSE}
miceOut <- readRDS(paste0(dataDir, "burn_data_mids_m25_it50_cart.rds"))
```

```{r}
miceOut$loggedEvents
```

It looks like many variables got dropped due to collinearity.

- The flags for missing patient/partner data seem to be especially susceptible.
- This isn't surprising since we have so much missing data and these flags have 
a lot of overlap.
- None of this should be a problem.

## Check Imputations

```{r, eval = FALSE}
pdf(paste0(plotDir, "mice_traceplots_m25_it50_cart.pdf"), onefile = TRUE)
plot(miceOut)
dev.off()
```

OK. After checking the plots in the PDF we just produced, it looks like the 
imputation models have all converged by 50 iterations. Let's see if the imputed 
values are plausible.

```{r, eval = FALSE}
### NOTE: This section is not currently working.
targets <- names(which(miceOut$method != ""))

pdf(paste0(plotDir, "mice_stripplots_m25_it50_cart.pdf"), onefile = TRUE)

v <- as.formula(targets[1])

for(v in targets)
  print(stripplot(miceOut, gender))
dev.off()
```

