---
title: "Prepare Data"
author: "Kyle M. Lang"
date: "2021-10-22"
output: 
 prettydoc::html_pretty:
    theme: architect
    highlight: github
    toc: true
    #toc_float: true
    toc_depth: 2
    #numbered_sections: true
    df_print: paged
    self_contained: false
    css: my_style.css
editor_options: 
  chunk_output_type: console
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      fig.align = "center",  
                      dev.args = list(bg = "transparent")
                      )
```

In this script, we'll explore/clean the burn data and prepare them for 
synthesis/imputation. We'll do the synthesis/imputation in a separate script.

---

# Data Ingest

---

First, we'll clear the workspace, load necessary packages, and define some 
global variables.

```{r, warning = FALSE, message = FALSE}
rm(list = ls(all = TRUE))

library(ggplot2)
library(plotly)
library(viridis)
library(haven)
library(mice)
library(xtable)
library(knitr)

dataDir <- "../../data/"
fn      <- "Social Impact Simulatie Totaal.sav"
```

Now, we'll read in the data.

```{r, cache = TRUE}
dat0 <- read_sav(paste0(dataDir, fn))

dim(dat0)
head(dat0)
```

That looks to have been successful. Let's explore the data.

---

# General Data Screening

---

First, we'll check the classes of each column.

```{r}
sapply(dat0, function(x) { 
  tmp <- class(x)
  tmp[length(tmp)] 
  }
  )
```

Everything read in as numeric. We need to convert the categorical variables to 
factors.

```{r}
facNames <- c("partner",
              "partner_incl",
              "gender",
              "Surgeries",
              "marital",
              "education",
              "dep_hist",
              "dep_treat",
              "gender_p",
              "education_p",
              "presence_p",
              "relation_t3")
dat0[facNames] <- lapply(dat0[facNames], as_factor)
```

Let's do a quick a dirty check for constant columns.

```{r}
nLevels <- sapply(dat0, function(x) length(unique(na.omit(x))))
any(nLevels == 1)
```

## Missing Data

How much missing data are we dealing with?

```{r}
pm <- colMeans(is.na(dat0))

range(pm[pm > 0])
hist(pm)
pm
```

What does the covariance coverage look like?

```{r}
cc <- md.pairs(dat0)$rr / nrow(dat0)

range(cc[cc < 1])

hist(cc)
```

Not awful. There is quite a large cluster of coverages around 30% - 40%, but we 
should be able to deal with that.

---

What kind of missing data patterns are we looking at?

```{r, out.width = "100%"}
## Calculate response patterns:
pats <- md.pattern(dat0)
```

Hmm...that's not a very enlightening figure. At least we can compute the number 
of unique response patterns.

```{r}
nrow(pats) - 1
```

---

# Exploring Partner Data

---

## Partner Flag Variables

What do the partner flags look like?

```{r, echo = FALSE}
t1 <- table(partner = dat0$partner, useNA = "ifany")
t2 <- xtabs(~ partner + partner_incl, data = dat0, addNA = TRUE)
kable(t1)
kable(t2)
```

I think the `partner` variable is junk.

- All cases with *unknown* partner status, and almost all cases who claim to 
have a partner are missing on the `partner_incl` variable.
   - I guess the first makes sense, but the second doesn't.
- Also, of the `r t1[3]` cases who claim not to have a partner, `r t2[3, 1]` 
simultaneously claim that their non-existant partner will participate.

## Partner Data

What about the partner data themselves?

```{r}
## Factor indicating partner participation:
pIncl <- dat0$partner_incl

## Extract the partner data:
pData <- dat0[grep("p$", colnames(dat0))]

## Separate demographics and scales:
scales     <- grepl("_t\\dp", colnames(pData))
pScaleData <- pData[scales]
pDemoData  <- pData[!scales]

## How much missing data on demos and scales?
colMeans(is.na(pDemoData))
colMeans(is.na(pScaleData))
```

We have a bit more missing data on the partner scales than partner demographics.

---

What about cases missing all demographics, all scales, or both?

```{r}
noPDemos  <- rowMeans(is.na(pDemoData)) == 1
noPScales <- rowMeans(is.na(pScaleData)) == 1
noPData   <- noPDemos & noPScales

noPDemosF <- factor(noPDemos, 
                    labels = paste("Partner Demos", c("Observed", "Missing"))
                    )
noPScalesF <- factor(noPScales, 
                     labels = paste("Partner Scales", c("Observed", "Missing"))
                     )
noPDataF <- factor(noPData, 
                   labels = paste("Partner Data", c("Observed", "Missing"))
                   )

kable(table(noPDemosF, noPScalesF))
kable(table(pIncl, noPDemosF, useNA = "ifany"))
kable(table(pIncl, noPScalesF, useNA = "ifany"))
kable(table(pIncl, noPDataF, useNA = "ifany"))
```

## Exploring/Fixing Strange Cases

All cases with missing values on `partner_incl` also miss all of their partner 
data.

- It seems pretty safe to assign all of these cases to the *partner does not 
participate* group.

```{r}
dat0$partner_incl[is.na(dat0$partner_incl)] <- "partner does not participate"
```

There is one case in the *partner does not participate* group who has partner 
data. What's going on there?

```{r}
flag <- dat0$partner_incl == "partner does not participate" & !noPData
sum(flag)

is.na(pData[flag, ])
```

This person has all partner data, so they should go in the *partner 
participates* group.

```{r}
dat0[flag, "partner_incl"] <- "partner participates"
```

Let's look at those crosstabs again.

```{r, echo = FALSE}
pIncl <- dat0$partner_incl

t1 <- table(noPDemosF, noPScalesF)
t2 <- table(pIncl, noPDemosF, useNA = "ifany")
t3 <- table(pIncl, noPScalesF, useNA = "ifany")
t4 <- table(pIncl, noPDataF, useNA = "ifany")
kable(t1)
kable(t2)
kable(t3)
kable(t4)
```

In the *partner participates* group, there are `r t3[1, 2]` cases with no scale 
data for partners and `r t4[1, 2]` cases with no partner data, at all.

- Should we move (on set of) these cases to the *partner does not participate* 
group?
- I guess that depends on how the participants ended up in the partner groups.
   - If they were assigned, we shouldn't move anyone (we should to an ITT 
   analysis).
   - If they simply reported, it should be safe to move them.

I don't think partner participation was randomly assigned, so let's move these 
cases, for now.

- Need to confirm this choice with Nancy et al.

```{r}
dat0[noPData, "partner_incl"] <- "partner does not participate"
```

```{r, echo = FALSE}
pIncl <- dat0$partner_incl

t1 <- table(pIncl, noPDemosF, useNA = "ifany")
t2 <- table(pIncl, noPScalesF, useNA = "ifany")
t3 <- table(pIncl, noPDataF, useNA = "ifany")
kable(t1)
kable(t2)
kable(t3)
```

---

# Exploring Patient Scale Data

---

## Missing Data

Let's see if any case is missing all of the patient scale data:

```{r}
scaleNames <- grep("_t\\d$", colnames(dat0), value = TRUE) %>% 
  grep("^rel", ., invert = TRUE, value = TRUE)

otherNames <- setdiff(colnames(dat0), 
                      c("ID_random", scaleNames, colnames(pData))
                      )

scaleData <- dat0[scaleNames]
otherData <- dat0[otherNames]

colPm <- colMeans(is.na(scaleData))
rowPm <- rowMeans(is.na(scaleData))

hist(colPm)

noScales <- rowPm == 1
sum(noScales)
mean(noScales)
```

The variablewise percents missing look pretty good. There are a few variables in 
the 40% to 50% missing range, but mostly the percents missing are around 25%.

The rowwise picture doesn't look so good, though. There are `r sum(noScales)` 
cases (`r round(100 * mean(noScales), 1)`%) with no scale data, at all.

## Evaluating Coverage

What kind of data *do* we have for the cases with entirely missing patient scale
data?

```{r}
noScalesF <- factor(noScales, 
                    labels = paste("Patient Scales", c("Observed", "Missing"))
                    )

t1 <- table(noScalesF)
t2 <- table(noScalesF, noPScalesF)
t3 <- table(noScalesF, noPDemosF)
t4 <- table(noScalesF, noPDataF)
t5 <- table(noScalesF, pIncl)

kable(t1)
kable(t2)
kable(t3)
kable(t4)
kable(t5)
```

Of the `r sum(noScales)` cases missing all patient scale data, `r t2[2, 2]` are 
also missing partner scale data and `r t3[2, 2]` are missing all partner data.

---

What kind of non-scale data do we have for the cases without patient scale data?

```{r, echo = FALSE, results = "hide"}
pats <- md.pattern(data.frame(otherData, pDemoData)[noScales, ], 
                   rotate.names = TRUE)
```

```{r, echo = FALSE}
pScales  <- ncol(scaleData)
pScalesP <- ncol(pScaleData)
pDemosP  <- ncol(pDemoData)
pOther   <- ncol(otherData)

pred <- pOther + pDemosP

pImp1 <- pScales + pScalesP + 2
pImp2 <- pImp1 + pDemosP
```

- None of these cases have data on the `relation_t3` variable or the 
`rel_duration_t3` variable. 
- As seen in the tables above, `r t3[2, 1]` cases have observed partner 
demographics.
- The response rates for the non-partner, non-scale variables looks pretty good, 
but these `r sum(noScales)` cases still face a brutal imputation challenge.
   - For the `r t3[2, 2]` cases with missing partner demographics, we're talking 
   about imputing `r pImp2` or `r pImp2 + 1` variables using only the 
   information in `r pOther - 1` or `r pOther` predictors.
   - For the `r t3[2, 1]` cases with observed partner demographics, the story 
   isn't much nicer. We still need to impute `r pImp1` or `r pImp1 + 1` 
   variables from `r pred - 4` or `r pred` predictors.

---

If we have to exclude the `r t1[2]` cases with completely missing patient scale 
data, we'll still have `r t1[1]` cases to work with. However, if we also have to 
exclude the cases with completely missing partner data, we'll only have 
`r t4[1, 1]` cases.

---

# Wrapping Up

---

Let's add some variables to flag the problematic cases discoverd above.
- We can try the analysis on reduced samples, if necessary.

```{r}
dat0$empty_patient_scales <- noScalesF
dat0$empty_partner_scales <- noPScalesF
dat0$empty_partner_data   <- noPDataF
```

*NOTE:* If we switch every case with entirely missing partner data to the 
*partner does not participate* group, then the `empty_partner_data` variable is 
perfectly redundant with the `partner_incl` variable.

---

I think we've got the data pretty much ready to go. Let's save the cleaned data.

- We'll do the data synthesis/imputation in another script.

```{r}
saveRDS(dat0, paste0(dataDir, "cleaned_burn_data.rds"))
```

---

END
